ms_0_intro

### 2018-02-07 man/ms_0_intro.Rmd
update

@@

### Sequías

Although summer drought is a characteristic feature of the Mediterranean climate [@Lionello2012], an increase in the frequency and severity of drought events have been recorded for the Mediterranean region [@Hoerling2011], particularly for southern Europe [@VicenteSerrano2014; @Spinoni2015; @Stagge2017] where a trend towards drier summers has been reported for the last three decades [@Spinoni2017b]. Además las proyecciones apuntan a que este incremento en la frecuencia continuará [@Spinoni2017a].

> notas sequia
> 
> * the drought events have been longest and most severe in the period 1991-2010 for mediterranean are of Southern Europe  
> * Seasonally, drought frequency is projected to increase everywhere in Europe for both scenarios in spring and summer, especially over southern Europe, and less intensely in autumn [@Spinoni2017a]
> Aumento de la frecuencia de sequías en spring and summer desde 1950 hasta 2014 [@Spinoni2017b]
> Both for frequency and severity, the evolution towards drier conditions is more relevant in the last three decades over Mediterranean area in summer, 
> an increase in the drought severity in the Iberian Peninsula has been observed in the last decades [@VicenteSerrano2014].

@@

* ojo hablar de que la sequía es un proceso natural, pero que se está viendo aumentado en los últimos años:

Drought is a natural phenomenon that occurs when water availability is significantly below normal levels over a long period and the supply cannot meet the existing demand.

> ver intro Ecosphere:
> Drought affects ecological systems across every climatic zone worldwide and is exacerbated by climate change and increasing anthropogenic water demands (Mishra and Singh 2010). Characterized by below-normal precipitation (Dai 2011), meteorological drought results from complex interactions between the atmosphere and hydrologic processes within the biosphere. Unlike aridity, which is a permanent feature of climate (Wilhite 1992), drought is a temporary extreme event (Palmer 1965, Mishra and Singh 2010) that can persist for extended time periods (months to years; Mishra and Singh 2010). Drought can cause significant changes in ecosystem productivity and water dynamics, and it is one of the most economically and ecologically disruptive extreme events affecting millions of people globally (Dai 2011).

Extreme climate events (e.g. droughts) severely affect forests and grasslands throuhg changes in plant physiology, phenology and carbon allocation [@Ummenhofer2017] (incluir citas para los efectos de la sequía sobre la vegetación).

@@

Climate change projections indicate that extreme events will become more common in the future (IPCC 2013), making it important that we understand how ecosystems respond to these events and the potential feedbacks to radiative forcing.

@@

>>>>

-- Sequías --

Drought is a natural phenomenon that occurs when water availability is significantly below-normal levels [@Dai2011]. Drought occurs over most part of the world and affects ecological systems across every climatic zone worldwide [@Mishra2010], being one of the most economically and ecologically disruptive extreme events affecting millions of people globally [@Dai2011]. Summer drought is a characteristic feature of the Mediterranean climate [@Lionello2012], but an increase in the frequency and severity of drought events have been recorded for the Mediterranean region [@Hoerling2011], particularly for southern Europe [@VicenteSerrano2014; @Spinoni2015; @Stagge2017] where a trend towards drier summers has been reported for the last three decades [@Spinoni2017b]. In addition, climate change projections suggest that extreme events, like droughts, will become more frequent and severe [@Trenberth2014; @IPCC2013] particularly for the Mediterranean area [@Spinoni2017a].

Drought can cause significant changes in ecosystem productivity and water dynamics, and extreme drought can severely affect forest through changes in plant physiology, phenology and carbon allocation (see @Ummenhofer2017 for a revision).

La vulnerabilidad de los bosques a los eventos climáticos extremos ...

no obstante los ecosistemas situados en zonas áridas responden mas rápidamente a la sequía.. ver VicenteSerrano2013

Existen evidencias que sugieren que muchos bosques son vulnerables a eventos climáticos extremos ... (Zhang) y esto puede ser especialmente relevante para especies situadas en el rear edge (completar) ...

... mejorar--> y esto puede ser especialmente importante en zonas como la región mediterránea,con un alto uso antrópico, donde la sequía presenta no solo efectos aislados, sino también la interacción con otros factores [@DoblasMiranda2017]. (ver abstract trabajo)

@@

### Sequías 
> notas sequia
> 
> * the drought events have been longest and most severe in the period 1991-2010 for mediterranean are of Southern Europe  
> * Seasonally, drought frequency is projected to increase everywhere in Europe for both scenarios in spring and summer, especially over southern Europe, and less intensely in autumn [@Spinoni2017a]
> Aumento de la frecuencia de sequías en spring and summer desde 1950 hasta 2014 [@Spinoni2017b]
> Both for frequency and severity, the evolution towards drier conditions is more relevant in the last three decades over Mediterranean area in summer, 
> an increase in the drought severity in the Iberian Peninsula has been observed in the last decades [@VicenteSerrano2014].

@@

@@

@@

.. ver [@VicenteSerrano2013] impacto de las sequías en diferentes zonas (i.e montañas)

Remote sensing can be used for studying droughts impacts on ecosystems [@Deshayes2006; @AghaKouchak2015; @Norman2016]

... ojo los remote sensing datos (aun con alta resolucion temporal y/o espacial) son insuficientes por si solos para demostrar 
 Remote sensing data, even high temporal or spatial resolution data, are insufficient alone to unequivocally demonstrate that drought is the causal agent of a particular change in reflectance values
